### 2018-05-16 Paper/examples.tex
..

\begin{table}[htbp]
\renewcommand*{\arraystretch}{1.2}
\begin{center}
\begin{tabular}{|c|c|}
\hline
Parameter & Distribution \\ \hline \hline
$r_w$ (Borehole radius, m) & $\mathcal{N}$(0.1,0.016) \\
$L$ (Borehole length, m) & $\mathcal{U}$[1120,1680] \\
$T_u$ (Transmissivity of upper aquifer, m$^2$/yr) & $\mathcal{U}$[63070,115600] \\
$H_u$ (Potentiometric head of upper aquifer, m) & $\mathcal{U}$[990,1110] \\
$T_l$ (Transmissivity of lower aquifer, m$^2$/yr) & $\mathcal{U}$[63.1,116] \\
$H_l$ (Potentiometric head of lower aquifer, m) & $\mathcal{U}$[700,820] \\
$K_w$ (Borehole hydraulic conductivity, m/yr) & $\mathcal{U}$[9855,12045] \\
\hline
\end{tabular}
\end{center}

\caption{Description and distributions of uncertain parameters in the borehole function
given by~\eqref{eq:bore}.}
\label{tab:bore}
\end{table}

Cheap function evaluations of the discharge $\mathcal{Q}(\vec{\theta})$ using
enables computing accurate estimates of $\mathcal{T}(\theta_i)$ through
sampling, with a sufficiently large Monte Carlo sample size.  Shown in
Figure~\ref{fig:sense_bore}~(left) are estimates of these indices corresponding to the
uncertain parameters in the borehole function using 10$^6$ pseudo-random
samples in the input parameter space. 
%\footnote{Although $\mathcal{T}(\theta_i)$ might converge with much fewer
%samples depending upon the model, we consider a large number that typically
%ensures a converged estimate for the purpose of illustration.} 
These estimates are used to verify fidelity of
parameter screening based on the methodology presented in Section~\ref{sec:method}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/sense_borehole}
  \includegraphics[width=0.51\textwidth]{./Figures/ub_conv_borehole}
\caption{
Left: Sobol' total sensitivity index for uncertain parameters in the borehole
function. Right: 
estimates of the screening parameter ($\widehat{\mathcal{C}_i\mu_i}$) is plotted
against number of samples considered for evaluating $\mu_i$. Corresponding estimates
for the converged Sobol' total effect index ($\mathcal{T}(\theta_i)$) in each case is also included in the
legend. }
\label{fig:sense_bore}
\end{center}
\end{figure}
In Figure~\ref{fig:sense_bore}~(right), we plot estimates of the screening parameter 
$\widehat{\mathcal{C}_i\mu_i}$ for a wide range of the number of 
samples used for approximating $\mu_i$ using~\eqref{eq:mu}.
Estimates for $\widehat{\mathcal{C}_i\mu_i}$ are found to be in excellent agreement
with $\mathcal{T}(\theta_i)$ even when small number of samples (5--10) are used. 
Consequently, the relative importance of uncertain 
parameters in the borehole function is found to be consistent 
with predictions based on the Sobol' index. 
In the considered intervals for the uncertain parameters, it is clear 
that the discharge is insensitive to $T_u$ and $T_l$. 
Moreover, the sensitivity to $K_w$ is also small. We exploit these findings to reduce
the dimensionality of the problem by 
discounting the uncertainties in $T_u$, $T_l$, and $K_w$ by fixing 
them at their respective nominal values.

Our goal as discussed is to gain computational advantage by constructing
surrogates in a reduced input parameter space. To this end, we use LAR to 
construct PCEs
in 5D and 4D spaces by fixing $\{T_u,T_l\}$ in the former and additionally
fixing $K_w$ in the latter at their respective mean values. In
Figure~\ref{fig:conv_bore}(left), we compare convergence of PCEs constructed in the
full space (7D) with those constructed in the two reduced spaces (4D and 5D)
using $\epsilon_{\mbox{\tiny{LOO}}}$ (Eq.~\ref{eq:loo}).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_borehole}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_borehole}
\caption{Left: A comparison of order of the leave-one-out-error 
($\epsilon_{\mbox{\tiny{LOO}}}$) as a function of number of regression samples
used for constructing the PCE in 4, 5, and 7 dimensions. Right: A comparison of
PDFs of the discharge, $\mathcal{Q}$, generated using 10$^{6}$ samples from
the marginal distributions of the uncertain parameters in each case.} 
\label{fig:conv_bore}
\end{center}
\end{figure}

As expected, it is observed that the PCE constructed in the 4D space converges
at a much faster rate. For instance, if a PCE with $\mathcal{O}(10^{-4})$
accuracy is sought, we need function evaluations at only about 50 sample points
in the 4D parameter space whereas the number of samples needed in the full 7D
space seems much higher. Latin hypercube sampling (LHS) was used in
each case. It must be pointed out that the error in Figure~\ref{fig:conv_bore}(left)
is not expected to decrease monotonically with the increase in sample size
owing to the penalty term in the regularized optimization problem in Eq.~\ref{eq:reg}.

@@

Figure~\ref{fig:conv_bore}(right) illustrates a comparison of the PDFs
of the
discharge, $\mathcal{Q}$ obtained by propagating 10$^6$ random
samples through the 7D PCE in the original input parameter domain as well as
the reduced order PCEs constructed in 4 and 5 dimensions. 
It is evident from this plot, that the PDFs agree quite favorably with each
other with respect to the modal estimate as well as the uncertainty associated with 
$\mathcal{Q}$. 
Consequently, it can be said that the reduced order PCE is verified in a
probabilistic sense. In other words, the mode as well as the uncertainty in the
observable is reliably captured and predicted by the reduced order PCE.

\subsection{Non-linear Oscillator}

@@

The above system has been studied extensively for its reliability using the following state function~\cite{Bucher:1989, 
Bucher:1990,Rajashekhar:1993,Schueremans:2005,Gayton:2003}:

\be
g(\bm{X}) = 3r - \left|z_\text{max}\right| = 3r - \left|\frac{2F}{m\omega_0^2}\sin\left(\frac{\omega_0t_1}{2}\right)\right|
\ee

>>>>

\begin{table}[htbp]
\renewcommand*{\arraystretch}{1.2}
\begin{center}
\begin{tabular}{|l||l|}
\hline
\textbf{Parameter Description} & \textbf{Distribution} \\ \hline \hline
Borehole radius, $r_w$ (m) & $\mathcal{N}$(0.1,0.016) \\
Borehole length, $L$ (m) & $\mathcal{U}$[1120,1680] \\
Transmissivity of upper aquifer, $T_u$ (m$^2$/yr) & $\mathcal{U}$[63070,115600] \\
Potentiometric head of upper aquifer, $H_u$ (m) & $\mathcal{U}$[990,1110] \\
Transmissivity of lower aquifer, $T_l$ (m$^2$/yr) & $\mathcal{U}$[63.1,116] \\
Potentiometric head of lower aquifer, $H_l$ (m) & $\mathcal{U}$[700,820] \\
Borehole hydraulic conductivity, $K_w$ (m/yr) & $\mathcal{U}$[9855,12045] \\
\hline
\end{tabular}
\end{center}

\caption{Description and distributions of uncertain parameters in the borehole function
given by~\eqref{eq:bore}. $\mathcal{N}(\mu,\sigma^2)$ denotes a normal distribution with
mean, $\mu$ and variance, $\sigma^2$. $\mathcal{U}[a,b]$ denotes a uniform distribution
with limits $a$ and $b$.}
\label{tab:bore}
\end{table}
%
Cheap function evaluations of the discharge $\mathcal{Q}(\vec{\theta})$ using
enables computing accurate estimates of $\mathcal{T}(\theta_i)$ through
sampling, with a sufficiently large Monte Carlo sample size.  Shown in
Figure~\ref{fig:sense_bore}~(left) are estimates of these indices corresponding to the
uncertain parameters in the borehole function using 10$^6$ pseudo-random
samples in the input parameter space. 
%\footnote{Although $\mathcal{T}(\theta_i)$ might converge with much fewer
%samples depending upon the model, we consider a large number that typically
%ensures a converged estimate for the purpose of illustration.} 
These estimates are used to verify fidelity of
parameter screening based on the methodology presented in Section~\ref{sec:method}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/sense_borehole}
  \includegraphics[width=0.51\textwidth]{./Figures/ub_conv_borehole}
\caption{
Left:Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the
borehole discharge function in~\eqref{eq:bore}. Right: 
Estimates of the screening parameter ($\widehat{\mathcal{C}_i\mu_i}$), plotted
against number of samples. Also included in the legend are estimates of $\mathcal{T}(\theta_i)$
in each case in the legend.}
\label{fig:sense_bore}
\end{center}
\end{figure}

In Figure~\ref{fig:sense_bore}~(right), we plot estimates of the screening parameter 
$\widehat{\mathcal{C}_i\mu_i}$ for a wide range of the number of 
samples used for approximating $\mu_i$ using~\eqref{eq:mu}.
Estimates for $\widehat{\mathcal{C}_i\mu_i}$ are found to be in excellent agreement
with $\mathcal{T}(\theta_i)$ even when small number of samples (5--10) are used. 
Consequently, the relative importance of uncertain 
parameters in the borehole function is found to be consistent 
with predictions based on the Sobol' index. 
In the considered intervals for the uncertain parameters, it is clear 
that the discharge is insensitive to $T_u$ and $T_l$. 
Moreover, the sensitivity to $K_w$ is also small. We exploit these findings to reduce
the dimensionality of the problem by 
discounting the uncertainties in $T_u$, $T_l$, and $K_w$ by fixing 
them at their respective nominal values.

Our goal as discussed is to gain computational advantage by constructing
surrogates in a reduced input parameter space. To this end, we use LAR to 
construct PCEs
in 5D and 4D spaces by fixing $\{T_u,T_l\}$ in the former and additionally
fixing $K_w$ in the latter at their respective mean values. In
Figure~\ref{fig:conv_bore}~(left), we compare convergence of PCEs constructed in the
full space (7D) with those constructed in the two reduced spaces (4D and 5D)
using $\epsilon_{\mbox{\tiny{LOO}}}$ (Eq.~\ref{eq:loo}).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_borehole}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_borehole}
\caption{Left: A comparison of order of the leave-one-out-error 
($\epsilon_{\mbox{\tiny{LOO}}}$) as a function of number of regression samples
used for constructing the PCE in 4, 5, and 7 dimensions. Right: A comparison of
PDFs of the discharge, $\mathcal{Q}$, generated using 10$^{6}$ samples from
the marginal distributions of the uncertain parameters in each case.} 
\label{fig:conv_bore}
\end{center}
\end{figure}
%
As expected, it is observed that the PCE constructed in the 4D space converges
at a much faster rate. For instance, if a PCE with $\mathcal{O}(10^{-4})$
accuracy is sought, we need function evaluations at only about 50 sample points
in the 4D parameter space whereas the number of samples needed in the full 7D
space seems much higher. Latin hypercube sampling (LHS) was used in
each case. It must be pointed out that the error in Figure~\ref{fig:conv_bore}~(left)
is not expected to decrease monotonically with the increase in sample size
owing to the penalty term in the regularized optimization problem in Eq.~\ref{eq:reg}.

@@

Figure~\ref{fig:conv_bore}~(right) illustrates a comparison of the PDFs
of the
discharge, $\mathcal{Q}$ obtained by propagating 10$^6$ random
samples through the 7D PCE in the original input parameter domain as well as
the reduced order PCEs constructed in 4 and 5 dimensions. 
It is evident from this plot, that the PDFs agree quite favorably with each
other with respect to the modal estimate as well as the uncertainty associated with 
$\mathcal{Q}$. 
Consequently, it can be said that the reduced order PCE is verified in a
probabilistic sense. In other words, the mode as well as the uncertainty in the
observable is reliably captured and predicted by the reduced order PCE. 
% 
\subsection{Non-linear Oscillator}

@@

The above system has been studied extensively for its reliability using
the following state function~\cite{Bucher:1989,Bucher:1990,Rajashekhar:1993,Schueremans:2005,Gayton:2003}:
%
\be
g(\bm{X}) = 3r - \left|z_\text{max}\right| = 3r - 
\left|\frac{2F}{m\omega_0^2}\sin\left(\frac{\omega_0t_1}{2}\right)\right|
\label{eq:limit}
\ee

\noindent where $z_\text{max}$ is the maximum displacement response of the system, $r$ denotes the displacement at
which either spring yields, $\bm{X}:~\{m,k_1,k_2,r,F,t_1\}$ is a set of uncertain parameters, and   
$\omega_0~=~\sqrt{\frac{k_1+k_2}{m}}$. The uncertain parameters are considered to be normally distributed with mean
and variance as provided below in Table~\ref{tab:osc}.

\begin{table}[htbp]
\renewcommand*{\arraystretch}{1.2}
\begin{center}
\begin{tabular}{|l||l|}
\hline
\textbf{Parameter Description} & \textbf{Distribution} \\ \hline \hline
Mass, $m$ (kg) & $\mathcal{N}$(1,0.05) \\
Time, $t_1$ (s) & $\mathcal{N}$(1,0.2) \\
Force, $F$ (N) & $\mathcal{N}$(1,0.2) \\
Displacement, $r$ (m) & $\mathcal{N}$(0.5,0.05) \\
Spring Constant, $k_1$ (Nm$^{-1}$) & $\mathcal{N}$(1,0.1) \\
Spring Constant, $k_2$ (Nm$^{-1}$) & $\mathcal{N}$(0.1,0.01) \\
\hline
\end{tabular}
\end{center}

\caption{Description and distributions of uncertain parameters in the limit state function
given by~\eqref{eq:limit}. $\mathcal{N}(\mu,\sigma^2)$ denotes a normal distribution with
mean, $\mu$ and variance, $\sigma^2$. $\mathcal{U}[a,b]$ denotes a uniform distribution
with limits $a$ and $b$.}
\label{tab:osc}
\end{table}

Converged estimates of $\mathcal{T}(\theta_i)$ are once again obtained easily by
evaluating $g(\bm{X})$ for a large set of samples in $\mathbb{R}^{|\bm{X}|}$ as
shown in Figure~\ref{fig:sense_osc}~(left). Estimates of the screening metric, 
$\widehat{\mathcal{C}_i\mu_i}$ are plotted against the number of samples in 
Figure~\ref{fig:sense_osc}~(right).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/sense_oscillator}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_oscillator}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the limit 
state function in~\eqref{eq:limit}. Right: 
Estimates of the screening parameter ($\widehat{\mathcal{C}_i\mu_i}$), plotted
against number of samples. Also included in the legend are estimates of $\mathcal{T}(\theta_i)$
in each case in the legend.}
\label{fig:sense_osc}
\end{center}
\end{figure}
